---
title: "Appendix"
author: "Game Estimations"
output:
  pdf_document: default
  html_document: default
---
### *Denote*:



 + *$P_j$*: is any player that is not $P_i$  


 + *Column $Sj$*: Range of possible strategies for Player j, which whenever summed with the other players' strategies, satisfy the given scenario. It can take any value in that range that is also 0, 1 or 2. For example, in *Table 3: Game Payoffs* we have 3 firms playing the game, where the second row has $Sj=[0,2]$, which is interpreted as the available strategies 0, 1 and 2 for all other players that are not Player i, such that whenever combined they satisfy $\sum_{i=1}^n S_i < K \leftrightarrow \sum_{i=1}^n S_i < 4$ (because K = number of firms + 1 and $Si=1$). Thus, we define Player 1's strategy as $S_1=1$, Player 2's strategy as $S_2=[0,2]$ and Player 3's strategy as $S_3=[0,2]$; then $(S_1, S_2, S_3) = \{(1,0,0), (1,1,0), (1,1,1), (1,0,1), (1,0,2), (1,2,0) \}$ are suggested in $Sj=[0,2]$ because they satisfy $S_1+S_2+S_3<4 \leftrightarrow \sum_{i=1}^nS_i < 4$.  


 + *Column $Vj(Sj, Si)$*: Range of possible payoffs for Player j.


&nbsp;
```{r setup, include=FALSE}
library(pander)
panderOptions('knitr.auto.asis', FALSE)
library(knitr)
library(kableExtra)
library(crayon)
```

```{r, include=FALSE}

###         Functions       ###
# By: Dayanara M. Diaz Vargas #
###                         ###
#---------------------------------------------------------------------------------------------
game_payoffs <- function(num_of_firms, p=0.5, K="none", S0=0, S1=1, S2=2, best_resp="Yes"){
  
  # set K
  if(K == "none"){
  K <- num_of_firms +1 # resource
  }
  
  
  # Probabilities 
  pr_e <- floor(num_of_firms/2)/num_of_firms
  pr_n <- 1-pr_e
  
  # Payoffs
  p0 <- S0
  p1 <- S1-p
  p2 <- S2-p
  p1_1 <- (S1-p)*pr_e+(-p)*pr_n
  p2_2 <- (S2-p)*pr_e+(-p)*pr_n
  
  # recurrent words
  vi <- "You get = "
  Si <- ", if you choose = "
  S_i <- " & the rest of the firms choose in aggregate = "
  
  # return
  
  if(best_resp == "Yes"){
      # also return payoffs
      payoffs <- data.frame( 
        Scenario = c("Sum_Si<K", "Sum_Si<K", "Sum_Si<K", "Sum_Si=>K", "Sum_Si=>K", "Sum_Si=>K"),
        Si = c(0, 1, 2, 1, 2, 0),
        Sj_left = c(S0, S0, S0, 
                    ifelse(((num_of_firms-2)*S2+S1*2)>=K, S1, S2), 
                    ifelse(((num_of_firms-2)*S2+S1*2)>=K, S1, S2), 
                    ifelse(((num_of_firms-2)*S2+S1*2)>=K, S1, S2)),
        Sj_right = c(S2, 
                     if((num_of_firms-2)*S0+S2+S1<K){
                       S2
                     } else if((num_of_firms-2)*S0+S1+S1<K){
                       S1 
                     } else {
                       S0
                     }, 
                     if((num_of_firms-2)*S0+S2+S2<K){
                       S2
                     } else if((num_of_firms-2)*S0+S1+S2<K){
                       S1
                     } else {
                       S0
                     }, 
                     S2, S2, S2),
        Si_Payoff = c(p0, p1, p2, p1_1, p2_2, p0),
        Sj_Payoff_left = c(p0, p0, p0, 
                           ifelse(((num_of_firms-2)*S2+S1*2)>=K, p1_1, p2_2),
                           ifelse(((num_of_firms-2)*S2+S1*2)>=K, p1_1, p2_2),
                           ifelse(((num_of_firms-2)*S2+S1*2)>=K, p1_1, p2_2)),
        Sj_Payoff_right = c(p2, 
                     if((num_of_firms-2)*S0+S2+S1<K){
                       p2
                     } else if((num_of_firms-2)*S0+S1+S1<K){
                       p1 
                     } else {
                       p0
                     }, 
                     if((num_of_firms-2)*S0+S2+S2<K){
                       p2
                     } else if((num_of_firms-2)*S0+S1+S2<K){
                       p1
                     } else {
                       p0
                     }, 
                     p2_2, p2_2, p2_2)
        )
      return(payoffs)
  } else {
    cat("Your payoff in each scenario, with ",  num_of_firms-1," other firms:\n")
     cat("\n")
      cat(vi , p0, Si, S0 , S_i, "[0, $\\infty$]\n")
      cat("\n")
        cat(vi, p1, Si, S1, S_i, "[0, ", K-S1-1, "]\n")
        cat("\n")
          cat(vi, p2, Si, S2, S_i, "[0, ", K-S2-1, "]\n")
          cat("\n")
            cat(vi, p1_1, Si, S1, S_i, " [", K-S1, " , $\\infty$]\n")
            cat("\n")
              cat(vi, p2_2, Si, S2, S_i,"[", K-S2, " , $\\infty$]\n")
                cat("\n")
                cat("\n")
                  cat("____________________________________________")
                  cat("\n")
                  cat("\n")
    }
}


#----------------------------------------------------------------------------------------------

game_pareto <- function(num_of_firms, K="none"){
  
  # set K
  if(K == "none"){
  K <- num_of_firms +1 # resource
  }
  
  # payoffs
    pays <- game_payoffs(num_of_firms=num_of_firms, best_resp="Yes")
    
    # pareto optimal strategy
    pareto <- pays[which((pays$Si_Payoff) == max(pays$Si_Payoff)), ]
      
      # number of other firms that choose each strategy 
                    nl <- 1 # num. of firms that choose the minimum strategy
                    nr <- 0 # num. of firms that choose the maximum strategy
                    nm <- 0 # num. of firms that choose the median strategy
                    m <- if(pareto$Sj_right==2 & pareto$Sj_left==0){
                            m <- 1
                            } else {
                             m <- "None"
                            }
                      
                    
                    if(m == "None"){
                                   nr <- ifelse(pareto$Sj_right==0, 
                                          floor((K - pareto$Si - nl*pareto$Sj_left)/1)-1, 
                                          floor((K - pareto$Si - nl*pareto$Sj_left)/pareto$Sj_right)-1)
                                   nl <- num_of_firms-nr-1
                                   
                      } else {while((nr*pareto$Sj_right+pareto$Si+nl*pareto$Sj_left+nm*m< K) & (nr+nl+nm+1<num_of_firms)){    
                            nm <- ifelse((nr*pareto$Sj_right+pareto$Si+nl*pareto$Sj_left+(nm+1)*m<K) & 
                                     (nr+nl+nm+1<num_of_firms) &
                                     (((ifelse(pareto$Sj_right==0, 
                                             floor((K - pareto$Si - nl*pareto$Sj_left - nm*m)/1), 
                                            floor((K - pareto$Si - nl*pareto$Sj_left - nm*m)/pareto$Sj_right)))*pareto$Sj_right+pareto$Si+nl*pareto$Sj_left+(nm)*m)<K),
                                   0, nm+1)
                                if((nr*pareto$Sj_right+pareto$Si+nl*pareto$Sj_left+nm*m<K) & (nr+nl+nm+1<num_of_firms)){
                                  nr <- ifelse(pareto$Sj_right==0, 
                                             floor((K - pareto$Si - nl*pareto$Sj_left - nm*m)/1), 
                                            floor((K - pareto$Si - nl*pareto$Sj_left - nm*m)/pareto$Sj_right))
                                    if((nr*pareto$Sj_right+pareto$Si+nl*pareto$Sj_left+nm*m< K) & (nr+nl+nm+1<num_of_firms)){
                                      nl <- num_of_firms-nr-nm-1
                                    }
                                }}
                        }
      
          # print result
            if(m == "None" & pareto$Sj_left==pareto$Sj_right){
                       cat("Pareto Optimality is reached when: $S_i$ = ", 
                          pareto$Si, " with $V_i(S_i, S_j)$ = ", 
                          pareto$Si_Payoff, ", if ", ifelse(pareto$Scenario=="Sum_Si<K", 
                          "$\\sum_{i=1}^n S_i < K$", "$\\sum_{i=1}^n S_i \\geq K$"), 
                          " & ", "$S_j$ = ", pareto$Sj_left,
                          " , with payoff $V_j(S_j, S_i)$ = ",
                          pareto$Sj_Payoff_left, ". With ",
                          nl, ifelse(nl==1, "other firm ", "other firms "), "choosing ", 
                          pareto$Sj_left," and viceversa.\n")
                       cat("\n")
                       cat("That is: \n")
                       if(pareto$Sj_left==pareto$Si){
                              cat(nl+1, " firms choose ", pareto$Sj_left, " units", ", each with a payoff of ", pareto$Sj_Payoff_left,".\n")
                       } else {
                              cat(nl, "firms choose ", pareto$Sj_left, " units, with a payoff of ", pareto$Sj_Payoff_left, ". \n", "1 firm chooses ", pareto$Si, " units, with a payoff of ", pareto$Si_Payoff, ". \n")
                       }
            } else if (m == "None" & pareto$Sj_left!=pareto$Sj_right){
                cat("Pareto Optimality is reached when: $S_i$ = ", 
                          pareto$Si, " with vi(Si, S_i) = ", 
                          pareto$Si_Payoff, ", if ", ifelse(pareto$Scenario=="Sum_Si<K", 
                          "$\\sum_{i=1}^n S_i < K$", "$\\sum_{i=1}^n S_i \\geq K$"),
                          " & ", "$S_j$ = [", pareto$Sj_left, " , ", pareto$Sj_right,
                          " ], with payoff $V_j(S_j, S_i)$ = [ ",
                          pareto$Sj_Payoff_left, " , ",
                          pareto$Sj_Payoff_right," ]. With ",
                          nl, ifelse(nl==1, "other firm choosing", "other firms choosing"), 
                          pareto$Sj_left, " and ", nr , 
                          ifelse(nr==1, "other firm choosing", "other firms choosing"), 
                          pareto$Sj_right," and viceversa.\n")
                       cat("\n")
                       cat("That is: \n")
                         if((pareto$Sj_left==pareto$Si) | (pareto$Sj_left==pareto$Si)){
                                cat(ifelse(pareto$Sj_left==pareto$Si, nl+1, nl), " firms choose ", pareto$Sj_left, " units", ", with a payoff of ", pareto$Sj_Payoff_left,
                                    ". \n", ifelse(pareto$Sj_left==pareto$Si, nr, nr+1), "firms choose ", pareto$Sj_right, " units, with a payoff of ", pareto$Sj_Payoff_right,
                                    ". \n")
                         } else {
                                cat(nl, "firms choose ", pareto$Sj_left, " units, with a payoff of ", pareto$Sj_Payoff_left, 
                                    ". \n", "1 firm chooses ", pareto$Si, " units, with a payoff of ", pareto$Si_Payoff, 
                                    ". \n", nr, "firms choose", pareto$Sj_right, "units, with a payoff of ", pareto$Sj_Payoff_right, 
                                    ". \n")
                         }    
            } else {
                        cat("Pareto Optimality is reached when: $S_i$ = ", 
                          pareto$Si, " with $V_i(S_i, S_j)$ = ", 
                          pareto$Si_Payoff, ", if ", ifelse(pareto$Scenario=="Sum_Si<K", 
                          "$\\sum_{i=1}^n S_i < K$", "$\\sum_{i=1}^n S_i \\geq K$"),
                          " & ", "$S_j$ = [", pareto$Sj_left, " , ",
                          pareto$Sj_right, " ], with payoff $V_j(S_j, S_i)$ = [",
                          pareto$Sj_Payoff_left, " , ",
                          pareto$Sj_Payoff_right, "]. With ",
                          nr, ifelse(nr==1, "other firm ", "other firms "), "choosing ",  pareto$Sj_right,
                          "; ", nm, ifelse(nm==1, "other firm choosing", "other firms choosing"), m,
                          ", and ", nl, ifelse(nl==1, "other firm", "other firms"),
                          " choosing", pareto$Sj_Payoff_left,", and viceversa.\n")
                          cat("\n")
                          cat("That is: \n")
                          if((pareto$Sj_left==pareto$Si) | (pareto$Sj_left==pareto$Si)){
                                cat(ifelse(pareto$Sj_left==pareto$Si, nl+1, nl), " firms choose ", pareto$Sj_left, " units", ", with a payoff of ", pareto$Sj_Payoff_left,
                                    ". \n", nm, " firms choose ", m, " units,  with a payoff of ", 0,
                                    ". \n", ifelse(pareto$Sj_left==pareto$Si, nr, nr+1), "firms choose ", pareto$Sj_right, " units, with a payoff of ", pareto$Sj_Payoff_right,
                                    ". \n")
                         } else {
                                cat(ifelse(pareto$Sj_left==pareto$Si, nl+1, nl), "firms choose ", pareto$Sj_left, " units, with a payoff of ", pareto$Sj_Payoff_left, 
                                    ". \n", nm, " firms choose ", m, " units,  with a payoff of ", 0,
                                    ". \n", ifelse(pareto$Sj_left==pareto$Si, nr, nr+1), "firms choose", pareto$Sj_right, "units, with a payoff of ", pareto$Sj_Payoff_right, 
                                    ". \n")
                         } 
            }
}

#-----------------------------------------------------------------------------------------------------

game_NE <- function(num_of_firms, K="none") {
  num_of_firms=3
  K="none"
  # set K
  if(K == "none"){
  K <- num_of_firms +1 # resource
  }
  
  # payoffs
    pays <- game_payoffs(num_of_firms=num_of_firms, best_resp="Yes")
    
  # Nash Equilibriums
    NEs <- pays[(pays$Si == pays$Sj_right | pays$Si == pays$Sj_right | (pays$Si == pays$Sj_right-pays$Sj_left-1 & pays$Si == 1)) & (pays$Scenario == "Sum_Si<K" |  pays$Scenario == "Sum_Si=>K"),]
  
    # results
    cat("There ", ifelse(nrow(NEs)==1, " is ", " are "), nrow(NEs), ifelse(nrow(NEs)==1, " Nash Equilibrium ", " Nash Equilibriums "), ".\n")
    cat("\n")
    for(i in 1:nrow(NEs)){
      cat("\n")
    cat(" ", "(", i, ")", "Nash Equilibrium ", i, ", for ", ifelse(NEs$Scenario[i]=="Sum_Si<K", 
                          "$\\sum_{i=1}^n S_i < K$", "$\\sum_{i=1}^n S_i \\geq K$"), ": \n")
    if(NEs$Sj_left[i]==NEs$Sj_right[i]){
      cat("  ", num_of_firms, " firms choose ", NEs$Sj_left[i], " units, each with a payoff of ", NEs$Sj_Payoff_left[i],". \n")
      cat("\n")
    } else if (NEs$Sj_left[i]!=NEs$Sj_right[i] & NEs$Sj_left[i]==NEs$Si[i]){
      cat("  ", num_of_firms, " firms choose ", NEs$Sj_left[i], " units, each with a payoff of ", NEs$Sj_Payoff_left[i],". \n")
      cat("\n")
    } else if (NEs$Sj_left[i]!=NEs$Sj_right[i] & NEs$Sj_right[i]==NEs$Si[i]){
      cat("  ", num_of_firms, " firms choose ", NEs$Sj_right[i], " units, each with a payoff of ", NEs$Sj_Payoff_right[i],". \n")
      cat("\n")
    } else {
      cat("  ", num_of_firms, " firms choose ", NEs$Si[i], " units, each with a payoff of ", NEs$Si_Payoff[i],". \n")
      cat("\n")
    }
    }
}
#-----------------------------------------------------------------------------------------------------





#-----------------------------------------------------------------------------------------------------
```

```{r, message = FALSE, echo= FALSE, warning = FALSE, results='asis'}


game_simulations <- function(num_simulations, best_resp="Yes", pareto="Yes", NE="Yes"){
  for (i in 2:num_simulations) {
    if(best_resp=="Yes" & pareto == "No"){
      cat("Game of ", i, " firms:")
      cat("\n")
      cat("Game of ", i, " firms:")
      cat("\n")
      print(game_payoffs(num_of_firms=i, best_resp=best_resp))
      cat("\n")
    } else if (best_resp=="Yes" & pareto == "Yes" & NE=="Yes") {
      cat("Game of ", i, " firms:")
      cat("\n")
      temp <- game_payoffs(num_of_firms=i, best_resp=best_resp)
      temp <- data.frame(lapply(temp[,-1], function(x) round(x, 3)))
      temp$Scenarios <- c('$ \\sum_{i=1}^n S_i < K$', '$ \\sum_{i=1}^n S_i < K$', '$ \\sum_{i=1}^n S_i < K$', '$ \\sum_{i=1}^n S_i \\geq K$', '$ \\sum_{i=1}^n S_i \\geq K$', '$ \\sum_{i=1}^n S_i \\geq K$')
      temp$Sj <- paste("[", temp$Sj_left, ",", temp$Sj_right, "]")
      temp$Sj_payoff <- paste("[", temp$Sj_Payoff_left, ",", temp$Sj_Payoff_right, "]")
      temp <- temp[, -c(2,3,5,6)]
      temp <- cbind(temp[,3], temp[,-3])
      colnames(temp) <- c("Scenarios", "Si", "Vi(Si, Sj)", paste0("Sj", footnote_marker_number(1)), paste0("Vj(Sj, Si)", footnote_marker_number(2)))
      print(kable_styling(kbl(temp, caption = "Game Payoffs", align = "l", booktabs = T, escape = F) %>%
  kable_paper(latex_options = c("striped"), full_width = F) %>%
  kable_styling(latex_options = "hold_position") %>%
  footnote(number = c("Sj = Range of possible strategies for Player j.", "Vj(Sj, Si) = Range of possible payoffs for Player j.")) %>%
  row_spec(0, bold = T)%>%
  column_spec(1, bold = T)))
      cat("\n")
      cat("---- Pareto Equilibrium --- ")
      cat("\n")
      game_pareto(num_of_firms=i)
      cat("\n")
      cat("---- Nash Equilibrium --- ")
      cat("\n")
      game_NE(num_of_firms=i)
      cat("\n\n\\pagebreak\n")
    } else {
    game_payoffs(num_of_firms=i, best_resp=best_resp)
    }
  }
}

```

## Game Predictions
```{r, message = FALSE, echo= FALSE, warning = FALSE, results='asis'}
#### run all
game_simulations(15)
```

## Reference for Students 
```{r, message = FALSE, echo= FALSE, warning = FALSE, results='asis'}
#### run all
game_simulations(15, best_resp="No")
```


